The title

### 2020-09-11 manuscript/manuscript.Rmd

---
title             : "The title"
shorttitle        : "Title"

author: 
  - name          : "First Author"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Writing - Original Draft Preparation
      - Writing - Review & Editing
  - name          : "Ernst-August Doelle"
    affiliation   : "1,2"
    role:
      - Writing - Review & Editing

affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
  - id            : "2"
    institution   : "Konstanz Business School"

@@

bibliography      : ["r-references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

csl               : "apa.csl"
documentclass     : "apa7"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
# Load packages
library(papaja)
library(reticulate)
library(MASS)
library(Matrix)
library(lme4)
library(buildmer)
library(emmeans)
library(magrittr)
library(tidyverse)

# Chunk options
knitr::opts_chunk$set(include = FALSE, echo = FALSE)

# Specify conda environment for Python
use_condaenv("r-reticulate", required = TRUE)

@@

```{r exp1-behavior}
# Define function for processing behavioral data
preproc_behav <- function(txt_fname){
  # Read log file
  dat <-  suppressMessages(suppressWarnings(read_tsv(txt_fname)))
  # Remove filler stimuli (well-known objects)
  dat %<>% filter(bek_unbek != "bekannt")
  # Add a column for the part of the experiment
  dat %<>% mutate(part = as_factor(Wdh - 210))
  # Assign stimuli to conditions based on keywords and button presses
  known <- dat %>% filter(part == "1" & Tastencode == 201) %>% pull(StimID)
  insight <- dat %>%filter(part == "2" & Bed == "richtig" & Tastencode %in% c(201, 202)) %>% pull(StimID)
  naive <- dat %>% filter(part == "2" & Bed == "falsch" & Tastencode %in% c(203, 204)) %>% pull(StimID)
  dat %<>% mutate(condition = as_factor(case_when(StimID %in% known ~ "Known",
                                                  StimID %in% insight ~ "Insight",
                                                  StimID %in% naive ~ "Naive")))
  # Remove unrealistically short RTs
  dat %<>% mutate(RT = replace(RT, RT < 200, NA))
  return(dat)}

# List of filenames for behavioral data
fnames_exp1 <- list.files("data/exp1/RT", pattern = ".txt", full.names = TRUE)

# Actually apply the preprocessing function
data_exp1 <- map_dfr(fnames_exp1, preproc_behav)

# Add factorized columns for participants and items (the random variables)
data_exp1 %<>% mutate(participant = as_factor(VPNummer), item = as_factor(StimID))

# Tabulate percentage of stimuli per condition
stims_exp1 <- table(data_exp1$participant, data_exp1$condition)
stims_exp1 <- colMeans(stims_exp1)
stims_exp1["Known"]/(3*120) # Three repititions and 120 objects
stims_exp1[c("Insight", "Naive")]/(3*60) # Three repitions and 60 objects each

# Select only relevant columns
data_exp1 %<>% select(part, condition, participant, item, RT)

# Tell MNE to exclude trials without a proper condition or RT
data_exp1 %<>% mutate(condition = replace(condition, condition == "Known" | is.na(RT), NA)) %>%
  mutate(condition = fct_explicit_na(condition, na_level = "Exclude")) %>% droplevels()
```

@@

# Define parameters for preprocessing
preproc_params = dict(n_components=15, method='fastica',           # ICA parameters
                      l_freq=0.1, h_freq=30,                       # Filter edges
                      event_id={'match': 221, 'mismatch': 222},    # EEG triggers
                      tmin=-0.5, tmax=1.498,                       # Length of epochs
                      baseline=(-0.2,0))                           # Baseline correction
reject = dict(eeg=200e-6)                                          # Rejection threshold

@@

# Define preprocessing function
def preproc(vhdr_fname, n_components, method, l_freq, h_freq, event_id, tmin, tmax, baseline):
  # Load EEG data
  raw = mne.io.read_raw_brainvision(vhdr_fname, preload=True)
  # Create new virtual EOG channels
  raw = mne.set_bipolar_reference(raw, 'Auge_u', 'Fp1', ch_name='VEOG', drop_refs=False)
  raw = mne.set_bipolar_reference(raw, 'F9', 'F10', ch_name='HEOG', drop_refs=False)
  raw.set_channel_types(mapping={'VEOG': 'eog', 'HEOG': 'eog'})
  # Add EasyCap electrode layout, removing any excessive channels
  montage = mne.channels.make_standard_montage('easycap-M1')
  raw.drop_channels(list(set(raw.ch_names) - set(montage.ch_names) - set(['VEOG', 'HEOG'])))
  raw.set_montage(montage=montage)
  # Re-reference to common average
  raw, _ = mne.set_eeg_reference(raw, 'average')
  # Run ICA on a copy of the data
  filt_raw = raw.copy()
  filt_raw.load_data().filter(l_freq=1, h_freq=None)
  ica = mne.preprocessing.ICA(n_components=n_components, random_state=12345, method=method)
  ica.fit(filt_raw)
  # Remove bad components based on correlations with EOG
  eog_indices, eog_scores = ica.find_bads_eog(raw)
  ica.exclude = eog_indices
  raw = ica.apply(raw)
  # Apply band-pass filter
  raw = raw.filter(l_freq=l_freq, h_freq=h_freq)
  # Epoching including baseline correction
  events, _ = mne.events_from_annotations(raw, verbose=False)
  epochs = mne.Epochs(raw, events=events, event_id=event_id, tmin=tmin, tmax=tmax,
                      baseline=baseline, preload=True)

  # # Autoreject                    
  # import autoreject
  # 
  # ar = autoreject.AutoReject(thresh_method='random_search', verbose=False)
  # epochs_clean = ar.fit_transform(epochs)
  # 
  # rejected = epochs_clean.drop_log
  # rejected[:] = [item for item in rejected if item != ['IGNORED']]
  # rejected = [i for i in range(len(rejected)) if rejected[i] != []]
  # 
  # ransac = autoreject.Ransac(verbose=False)
  # epochs = ransac.fit_transform(epochs)
  # 
  # reject = autoreject.get_rejection_threshold(epochs)

  return epochs

# List of raw EEG filenames
fnames_exp1 = sorted(glob.glob('data/exp1/EEG/*.vhdr'))

# Check if preprocessing was done already (delete the file to re-run)
if os.path.exists('output/exp1-epo.fif'):
  # Load preprocessed data from file
  epochs_exp1 = mne.read_epochs('output/exp1-epo.fif', preload=True)
else:
  # Actually apply the preprocessing function
  epochs_exp1 = [preproc(fname, **preproc_params) for fname in fnames_exp1]
  # Combine into a single data set
  epochs_exp1 = mne.concatenate_epochs(epochs_exp1)
  # Add metadata from R
  epochs_exp1.metadata = r.data_exp1
  # Backup epochs to the export folder
  epochs_exp1.save('output/exp1-epo.fif')

# Remove trials with invalid condition or RTs
epochs_exp1 = epochs_exp1['condition != "Exclude"']

# # Compute evoked potentials for each part x condition
# evokeds_exp1 = dict()
# for pt in ['1', '2', '3']:
#   for cn in ['Insight', 'Naive']:
#     query = 'part == "' + pt + '" and condition == "' + cn + '"'
#     evokeds_exp1['Part' + pt + '_' + cn] = epochs_exp1[query].average()
# mne.viz.plot_compare_evokeds(evokeds_exp1, cmap=('condition', 'viridis'), picks='P8')

# # Get indices of bad epochs despite ICA
# rej_exp1 = epochs_exp1.copy().drop_bad(reject=reject).drop_log
# rej_exp1[:] = [item for item in rej_exp1 if item != ['IGNORED']]
# rej_exp1 = [i for i in range(len(rej_exp1)) if rej_exp1[i] != []]
```

```{r exp1-analysis}

# Import epochs from Python, converting Volts to Microvolts
epochs_exp1 <- py$epochs_exp1$get_data()*1e6

# Set rejected epochs to NA
rej_exp1 <- py$rej_exp1 + 1
epochs_exp1[rej_exp1,,] <- NA

@@

# Define function to compute single-trial ERPs
compute_erps <- function(epochs, els, name, start, stop, roi){
  erps <- epochs[, which(els %in% roi), which(times %in% start:stop)]
  erps <- apply(erps, 1, mean, na.rm = TRUE)
  return(erps)}

# Apply the function and bind ERPs to the behavioral data
data_exp1 <- pmap_dfc(comps, compute_erps, epochs = epochs_exp1, els = els) %>% set_names(comps$name) %>% cbind(data_exp1, .)

# Remove trials with missing ERPs or RTs
rej_exp1_erp <- data_exp1 %>% filter(is.na(P1)) %>% count(participant, .drop = FALSE) %>% pull(n)
data_exp1 %<>% filter(!is.na(P1))
rej_exp1_rt <- data_exp1 %>% filter(is.na(RT)) %>% count(participant, .drop = FALSE) %>% pull(n)
data_exp1 %<>% filter(!is.na(RT))

# Remove any trials that don't beling to either of the two conditions
data_exp1 %<>% filter(condition != "Known") %>% droplevels()

# Set sliding difference coding for condition (insight-naive) and part (2-1, 3-1)
t(contrasts.condition <- t(cbind(c("Insight" = 1, "Naive" = -1))))
t(contrasts.part <- t(cbind(c("1" = -1, "2" = 1, "3" = 0),
                            c("1" = 0, "2" = -1, "3" = 1))))
contrasts(data_exp1$condition) <- ginv(contrasts.condition)
contrasts(data_exp1$part) <- ginv(contrasts.part)

# Parameters for linear mixed-effects models (LMMs)
form <- tabulate.formula(~ part*condition + (part*condition|participant) + (part*condition|item))
form %<>% mutate(block = replace(block, is.na(grouping), "fixed"))
ctrl_params <- lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))

# Parameters for follow-up contrasts
specs = pairwise ~ condition|part

# Define function to compute LMMs and contrasts
compute_models <- function(dep, formula, data, control, specs){
  build <- buildmer(dep = dep, formula = formula, data = data, ddf = "Satterthwaite", calc.anova = TRUE, control = control)
  emm_options(lmer.df = "Satterthwaite", lmerTest.limit = Inf)
  conts <- emmeans(build@model, specs = specs, infer = TRUE)
  mod <- list("model" = build@model, "anova" = build@anova, "summary" = build@summary, "contrasts" = conts)
  return(mod)}

# Check if models were computed already (delete the file to re-run)
if (file.exists("output/exp1-stats.RDS")){
   # Load models from file
  models_exp1 <- readRDS("output/exp1-stats.RDS")
} else {
  # Actually apply the modelling function
  models_exp1 <- map(comps$name, compute_models, formula = form, data = data_exp1, control = ctrl_params, specs = specs) %>%
    set_names(comps$name)
  # Backup epochs to the export folder
  saveRDS(models_exp1, "output/exp1-stats.RDS")}
```

```{python exp2-preprocessing}
# List of raw EEG filenames
fnames_exp2 = sorted(glob.glob('data/exp2/EEG/*.vhdr'))

# Check if preprocessing was done already (delete the file to re-run)
if os.path.exists('output/exp2-epo.fif'):
  # Load preprocessed data from file
  epochs_exp2 = mne.read_epochs('output/exp2-epo.fif', preload=True)
else:
  # Actually apply the preprocessing function
  epochs_exp2 = [preproc(fname, **preproc_params) for fname in fnames_exp2]
  # Combine into a single data set
  epochs_exp2 = mne.concatenate_epochs(epochs_exp2)
  # Backup epochs to the export folder
  epochs_exp2.save('output/exp2-epo.fif')

# Get indices of bad epochs despite ICA
rej_exp2 = epochs_exp2.copy().drop_bad(reject=reject).drop_log
rej_exp2[:] = [item for item in rej_exp2 if item != ['IGNORED']]
rej_exp2 = [i for i in range(len(rej_exp2)) if rej_exp2[i] != []]
```

```{r exp2-analysis}
# List of filenames for behavioral data
fnames_exp2 <- list.files("data/exp2/RT", pattern = ".txt", full.names = TRUE)

# Actually apply the preprocessing function
data_exp2 <- map_dfr(fnames_exp2, preproc_behav)

# Add factorized columns for participants and items (the random variables)
data_exp2 %<>% mutate(participant = as_factor(VPNummer), item = as_factor(StimID))

# Tabulate percentage of stimuli per condition
stims_exp2 <- table(data_exp2$participant, data_exp2$condition)
stims_exp2 <- colMeans(stims_exp2)
stims_exp2["Known"]/(3*120) # Three repititions and 120 objects
stims_exp2[c("Insight", "Naive")]/(3*60) # Three repitions and 60 objects each

# Import epochs from Python, converting Volts to Microvolts
epochs_exp2 <- py$epochs_exp2$get_data()*1e6

# Set rejected epochs to NA
rej_exp2 <- py$rej_exp2 + 1
epochs_exp2[rej_exp2,,] <- NA

# Apply function to compute single-trial ERPs and to the behavioral data
data_exp2 <- pmap_dfc(comps, compute_erps, epochs = epochs_exp2, els = els) %>% set_names(comps$name) %>% cbind(data_exp2, .)

# Remove trials with missing ERPs or RTs
rej_exp2_erp <- data_exp2 %>% filter(is.na(P1)) %>% count(participant, .drop = FALSE) %>% pull(n)
data_exp2 %<>% filter(!is.na(P1))
rej_exp2_rt <- data_exp2 %>% filter(is.na(RT)) %>% count(participant, .drop = FALSE) %>% pull(n)
data_exp2 %<>% filter(!is.na(RT))

# Remove any trials that don't beling to either of the two conditions
data_exp2 %<>% filter(condition != "Known") %>% droplevels()

# Set sliding difference coding for condition (insight-naive) and part (2-1, 3-1)
contrasts(data_exp2$condition) <- ginv(contrasts.condition)
contrasts(data_exp2$part) <- ginv(contrasts.part)

# Check if models were computed already (delete the file to re-run)
if (file.exists("output/exp2-stats.RDS")){
   # Load models from file
  models_exp2 <- readRDS("output/exp2-stats.RDS")
} else {
  # Actually apply the modelling function
  models_exp2 <- map(comps$name, compute_models, formula = form, data = data_exp2, control = ctrl_params, specs = specs) %>%
    set_names(comps$name)
  # Backup epochs to the export folder
  saveRDS(models_exp2, "output/exp2-stats.RDS")}
```

>>>>

---
title             : "Title of the Thesis Goes Here"
shorttitle        : "Short title goes here"

author: 
  - name          : "Alexander Enge"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Rudower Chaussee 18, 12489 Berlin"
    email         : "alexander.enge@hu-berlin.de"

affiliation:
  - id            : "1"
    institution   : "Humboldt-Universität zu Berlin"

@@

bibliography      : "r-references.bib"
zotero            : "aha"

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

csl               : "apa.csl"
documentclass     : "apa7"
classoption       : "man,11pt"
output:
  papaja::apa6_pdf:
    latex_engine:   "xelatex"

header-includes:
  - \usepackage{setspace}
  - \AtBeginEnvironment{tabular}{\doublespacing}
  - \AtBeginEnvironment{lltable}{\doublespacing}
  - \AtBeginEnvironment{tablenotes}{\doublespacing}
  - \captionsetup[table]{font={stretch=1.5}}
  - \captionsetup[figure]{font={stretch=1,scriptsize}}
  - \raggedbottom
---

```{r setup, include=FALSE}
# Load packages
library(papaja)
library(reticulate)
library(lme4)
library(buildmer)
library(tidyverse)
library(magrittr)
library(cowplot)

# Chunk options
knitr::opts_chunk$set(include = FALSE, fig.align = "center", fig.out = "100%", fig.width = 10)

# Specify Python environment
renv::use_python(type = "virtualenv")

@@

Introduction section goes here.

# Experiment 1

```{r exp1-preparation}
# Define function for reading and formatting behavioral data
read_behav <- function(txt_fname){
  # Read log file
  dat <- suppressMessages(suppressWarnings(read_tsv(txt_fname)))
  # Rename some columns
  dat %<>% rename(participant = VPNummer, item = StimID) %>%
    # Add a column for the part of the experiment
    mutate(part = case_when(Wdh %in% c(211, 231) ~ "I",
                            Wdh %in% c(212, 232) ~ "II",
                            Wdh %in% c(213, 233) ~ "III",
                            Wdh == 234 ~ "IV") %>% as_factor()) %>%
    # Remove filler stimuli (i.e. well-known objects)
    filter(bek_unbek != "bekannt")
  # Assign stimuli to conditions based on keywords and button presses
  excl_known <- dat %>% filter(part == "I" & Tastencode %in% c(201, 251)) %>% pull(item)
  excl_insight <- dat %>% filter(part == "II" & grepl("richtig", Bed) & !Tastencode %in% c(201, 202, 251, 252)) %>% pull(item)
  excl_naive <- dat %>% filter(part == "II" & grepl("falsch", Bed) & !Tastencode %in% c(203, 204, 253, 254)) %>% pull(item)
  insight <- dat %>% filter(part == "II" & grepl("richtig", Bed) & Tastencode %in% c(201, 202, 251, 252)) %>% pull(item)
  naive <- dat %>% filter(part == "II" & grepl("falsch", Bed) & Tastencode %in% c(203, 204, 253, 254)) %>% pull(item)
  # Create a new condition column, excluding trials with known stimuli or without condition
  dat %<>% mutate(condition = case_when(item %in% excl_known ~ "Excl_known",
                                        item %in% excl_insight ~ "Excl_insight",
                                        item %in% excl_naive ~ "Excl_naive",
                                        item %in% insight ~ "Insight",
                                        item %in% naive ~ "Naive") %>%
                    factor(levels = c("Insight", "Naive", "Excl_insight", "Excl_naive", "Excl_known")))
  # In Experiment 3, we need an additional column for position (upright vs. inverted)
  if("richtig_inv" %in% dat$Bed){
    dat %<>% mutate(position = as_factor(ifelse(grepl("inv", Bed), "Inverted", "Upright"))) %>%
      # Return only relevant columns
      select(part, position, condition, participant, item, RT)
  } else {
    dat %<>% select(part, condition, participant, item, RT)}
  return(dat)}

# Read behavioral data for Experiment 1
fnames_exp1 <- list.files("data/exp1/RT", pattern = ".txt", full.names = TRUE)
data_exp1 <- map(fnames_exp1, read_behav)

# Check average number of stimuli per condition
(stims_exp1 <- map(data_exp1, function(x){table(x$condition)/3}) %>% bind_rows() %>% colMeans())
```

@@

# Set parameters for EEG preprocessing
preproc_params = dict(n_components=15, method='fastica',           # ICA parameters
                      l_freq=0.1, h_freq=30,                       # Filter edges
                      event_id={'match': 221, 'mismatch': 222},    # EEG triggers
                      tmin=-0.5, tmax=1.498,                       # Length of epochs
                      baseline=(-0.2,0),                           # Baseline correction
                      reject=dict(eeg=200e-6))                     # Rejection threshold

@@

# Define function for preprocessing EEG data
def preproc(vhdr_fname, metadata, n_components, method, l_freq, h_freq, event_id, tmin, tmax, baseline, reject):
  # Read EEG data
  raw = mne.io.read_raw_brainvision(vhdr_fname, preload=True)
  # Create virtual EOG channels
  raw = mne.set_bipolar_reference(raw, 'Auge_u', 'Fp1', ch_name='VEOG', drop_refs=False)
  raw = mne.set_bipolar_reference(raw, 'F9', 'F10', ch_name='HEOG', drop_refs=False)
  raw.set_channel_types(mapping={'VEOG': 'eog', 'HEOG': 'eog'})
  # Add EasyCap electrode layout, removing any excessive channels
  montage = mne.channels.make_standard_montage('easycap-M1')
  raw.drop_channels(list(set(raw.ch_names) - set(montage.ch_names) - set(['VEOG', 'HEOG'])))
  raw.set_montage(montage=montage)
  # Re-reference to common average
  raw, _ = mne.set_eeg_reference(raw, 'average')
  # Run ICA on a copy of the data
  filt_raw = raw.copy()
  filt_raw.load_data().filter(l_freq=1, h_freq=None)
  ica = mne.preprocessing.ICA(n_components=n_components, random_state=12345, method=method)
  ica.fit(filt_raw)
  # Remove bad components based on correlations with EOG
  eog_indices, eog_scores = ica.find_bads_eog(raw)
  ica.exclude = eog_indices
  raw = ica.apply(raw)
  # Apply band-pass filter
  raw = raw.filter(l_freq=l_freq, h_freq=h_freq)
  # Epoching including baseline correction
  events, _ = mne.events_from_annotations(raw, verbose=False)
  epochs = mne.Epochs(raw, events=events, event_id=event_id, tmin=tmin, tmax=tmax, baseline=baseline, preload=True)
  # Add behavioral data
  epochs.metadata = metadata
  # Reject bad epochs
  epochs = epochs.drop_bad(reject=reject)
  return epochs

# List raw EEG filenames
fnames_exp1 = sorted(glob.glob('data/exp1/EEG/*.vhdr'))

# Import behavioral data from R
metadata_exp1 = r.data_exp1

# Check if preprocessing was done already (delete the file to re-run)
if os.path.exists('output/exp1-epo.fif'):
  # Read preprocessed data from file
  epochs_exp1 = mne.read_epochs('output/exp1-epo.fif', preload=True)
else:
  # Preprocess EEG data for Experiment 1
  epochs_exp14 = [preproc(vhdr_fname=fname, metadata=meta, **preproc_params) for fname, meta in zip(fnames_exp1, metadata_exp1)]
  # Combine epochs into a single data set
  epochs_exp1 = mne.concatenate_epochs(epochs_exp1)
  # Backup epochs to the output folder
  epochs_exp1.save('output/exp1-epo.fif')

# # Get indices of bad epochs despite ICA
# rej_exp1 = epochs_exp1.drop_log
# rej_exp1[:] = [item for item in rej_exp1 if item != ['IGNORED']]
# rej_exp1 = [i for i in range(len(rej_exp1)) if rej_exp1[i] != []]

# Define function to compute grand averages per condition
def compute_evokeds(epochs):
  evokeds = dict()
  evokeds_dat = dict()
  # Each part becomes a dictionary of conditions
  for pt in ['I', 'II', 'III']:
    evokeds[pt] = dict()
    evokeds_dat[pt] = dict()
    # Each condition becomes a list of participants
    for cn in ['Insight', 'Naive']:
      evokeds[pt][cn] = list()
      evokeds_dat[pt][cn] = list()
      # For every participant we average trials separately for all parts and conditions
      for vp in epochs.metadata['participant'].unique():
        query = 'participant == "' + vp + '" and part == "' + pt + '" and condition == "' + cn + '"'
        evokeds[pt][cn].append(epochs[query].average())
      # Compute grand averages across participants for this part and condition
      evokeds[pt][cn] = mne.grand_average(evokeds[pt][cn])
      # Export only the actual data for R
      evokeds_dat[pt][cn] = evokeds[pt][cn].data
  return(evokeds, evokeds_dat)

# Compute grand averages for Experiment 1
evokeds_exp1, evokeds_dat_exp1 = compute_evokeds(epochs=epochs_exp1)

# # Plot evoked potentials
# import matplotlib.pyplot as plt
# mne.viz.plot_compare_evokeds(evokeds=evokeds_exp1['II'], picks='P8'); plt.show()
# evokeds_exp1['II']['Insight'].data = evokeds_exp1['II']['Insight'].data - evokeds_exp1['II']['Naive'].data
# evokeds_exp1['II']['Insight'].plot_topomap(times=[0.125, 0.175, 0.5], average=0.05, vmin=-1, vmax=1, cmap='viridis'); plt.show()
```

```{r exp1-analysis}
# Re-import behavioral data
data_exp1 <- py$epochs_exp1$metadata

# Factorize some columns
data_exp1 %<>% mutate(part = factor(part, levels = c("I", "II", "III")),
                      condition = factor(condition, levels = c("Insight", "Naive", "Excl_insight", "Excl_naive", "Excl_known")),
                      participant = factor(participant),
                      item = factor(item))

# Check number of rejected epochs
rejected_exp1 <- data_exp1 %>% group_by(participant) %>% tally() %>% mutate(n = 360 - n) %>% pull(n)
mean(rejected_exp1); median(rejected_exp1); range(rejected_exp1)

@@

# Define function to compute single-trial mean ERP amplitudes
compute_erps <- function(epochs, els, name, start, stop, roi){
  erps <- epochs[, which(els %in% roi), which(times %in% start:stop)]
  erps <- apply(erps, 1, mean, na.rm = TRUE)
  return(erps)}

# Check if single-trial ERPs were computed already (delete the file to re-run)
if (file.exists("output/exp1-erps.RDS")){
  # Read data from file
  data_exp1 <- readRDS("output/exp1-erps.RDS")
} else {
  # Import epochs from Python, converting Volts to Microvolts
  epochs_exp1 <- py$epochs_exp1$get_data()*1e6
  # Compute single-trial ERPs for Experiment 1
  data_exp1 <- pmap_dfc(comps, compute_erps, epochs = epochs_exp1, els = els) %>%
    set_names(comps$name) %>% cbind(data_exp1, .) %T>% saveRDS("output/exp1-erps.RDS")}

# Remove any trials excluded from conditions
data_exp1 %<>% filter(condition %in% c("Insight", "Naive")) %>% droplevels()

# Contrast coding for condition (insight-naive) and part (2-1, 3-1)
contrasts_condition <- t(cbind(c("Insight" = 1, "Naive" = -1)))
contrasts_part <- t(cbind(c("I" = -1, "II" = 1, "III" = 0), c("I" = 0, "II" = -1, "III" = 1)))
contrasts(data_exp1$condition) <- MASS::ginv(contrasts_condition)
contrasts(data_exp1$part) <- MASS::ginv(contrasts_part)

# Set formula and parameters for linear mixed-effects regression models (LMMs)
form_exp12 <- tabulate.formula(~ part*condition + (part*condition|participant) + (part*condition|item))
form_exp12 %<>% mutate(block = replace(block, is.na(grouping), "fixed"))
ctrl_params <- lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))

# Set formula for follow-up contrasts
specs_exp12 = pairwise ~ condition|part

# Define function to compute LMMs and follow-up contrasts for each component
compute_models <- function(dep, formula, data, control, specs){
  require(emmeans)
  emm_options(lmer.df = "Satterthwaite", lmerTest.limit = Inf)
  build <- buildmer(dep = dep, formula = formula, data = data, control = control, REML = FALSE,
                    ddf = "Satterthwaite", calc.anova = TRUE)
  conts <- emmeans(build@model, specs = specs, infer = TRUE, data = data)$contrasts %>% as.data.frame()
  mod <- list("model" = build@model, "anova" = build@anova, "summary" = build@summary, "contrasts" = conts)
  return(mod)}

# Check if models were computed already (delete the file to re-run)
if (file.exists("output/exp1-stats.RDS")){
  # Read models from file
  models_exp1 <- readRDS("output/exp1-stats.RDS")
} else {
  # Compute LMMs for Experiment 1
  models_exp1 <- map(comps$name, compute_models, formula = form_exp12, data = data_exp1, control = ctrl_params,
                     specs = specs_exp12) %>% set_names(comps$name) %T>% saveRDS("output/exp1-stats.RDS")}
```

## Methods

### Participants

Participants for Experiment 1 were 24 German native speakers (13 female, 11 male) with a mean age of 24 years (range 18 to 31) and no history psychological disorder or treatment. All participants were right-handed according to the Edinburgh inventory [@oldfield1971] and reported normal or corrected-to-normal vision. They gave written informed consent before starting the experiment and received a compensation of €8 per hour for participating. All experiments subsequently reported were carried out in accordance with the Declaration of Helsinki and approved by the local ethics committee.

### Materials and Procedure

Stimuli for Experiments 1 and 2 consisted of 240 grayscale photographs of real-world objects, 120 of which were well-known everyday objects (e.g. a bicycle, a toothbrush), whereas the other 120 were rare objects presumed to be unfamiliar for the majority of participants (e.g. a galvanometer, an udu drum). A list of all objects can be found in Appendix A. They were presented on a light blue background with a size of 207 × 207 pixels on a 19-inch LCD monitor with a resolution of 1,280 × 1,024 pixels and a refresh rate of 75 Hz. At a standardized viewing distance of 90 cm, the images of the objects subtended approximately 3.9 degress of participants' horizontal and vertical visual angle.

For each unfamiliar object, a noun--verb pair was selected describing the object's typical function or use in a way that could typically be related to its visual features and their configuration (e.g. current--measuring, pottery--drumming). In the *insight* condition, the presentation of each unfamiliar object was preceded by its respective description, whereas in the *naive* condition, it was preceded by a non-matching description belonging to one of the other objects. In both conditions, the noun and the verb were presented one above the other in black letters on a light blue background. For each participant, half of the unfamiliar objects were presented in the insight condition (i.e. with a matching description) and the other half were presented in the naive condition (i.e. with a non-matching description). All participants saw each unfamiliar object in only one of the two conditions and the assignment of objects to either of the two conditions was counterbalanced across participants. The experiment was programmed and displayed using Presentation® software (Neurobehavioral Systems, Inc., Berkeley, CA, www.neurobs.com).

Each experimental session consisted of three parts (Figure \@ref(fig:exp1-plot)A). In Part I, after written informed consent was obtained and the EEG was prepared, all 240 familiar and unfamiliar objects were presented once in random order and without verbal descriptions. Each trial consisted of a fixation cross presented in the middle of the screen for 0.5 s, followed by the presentation of the object until participants made a response or until a time out after 3 s. The intertrial interval until the presentation of the next fixation cross was 0.5 s and participants took a self-timed break after each block of 60 objects. The task, which was kept the same throughout all three parts and experiments, was to classify each object using one of four response alternatives: (a) "I know what this is or have a strong assumption," (b) "I have a strong assumption," (c) "I have rather no assumption what this is," or (d) "I do not know what this is and have no assumotion." Participants were asked to respond as quickly and as accurately as possible by pressing one out of four buttons with the index or middle finger of their left or right hand, respectively. The mapping of the rating scale to the four buttons (left to right or right to left) was counterbalanced across participants. Because we were interested in the acquisition of semantic knowledge about novel objects only, all objects for which the participant indicated knowing the object in Part I were excluded from further analysis (`r scales::percent(stims_exp1["Excl_known"]/120, accuracy = 0.1)` of rare objects per participant).

(ref:figure-1-caption) Procedure and results of Experiment 1. (A) In Part I, participants were presented with 120 unfamiliar objects and indicated whether they knew what kind of object they were viewing. In Part II, half of the objects were preceded by a matching description (in purple color for illustration), leading to semantic insight, and the other half with a non-matching description (in petrol color for illustration), leading to naive perception. In Part III, the same objects were presented again without the descriptions. (B) Bar plots and (C) ERP waveforms and scalp topographies separately for objects for which participants experienced semantic insight and naive perception in Parts I, II, and III. Semantic insight was associated with significantly more negative amplitudes in the N1 component in Part II, significantly less negative amplitudes in the N400 component in Parts II and III, and significantly more positive amplitudes in the P1 component in Part III. Error bars show standard errors of the mean. \newline\**p* \< .05. \*\**p* \< .01. \*\*\**p* \< .001.

```{r exp1-plot, include=TRUE, fig.height=12, fig.cap = "(ref:figure-1-caption)"}
# Manually extract significance levels from contrasts for plotting
stars_exp1 = list(P1 = c(I = NA, II = NA, III = "*"),
                  N1 = c(I = NA, II = "**", III = NA),
                  N400 = c(I = NA, II = "***", III = "**"))

# Import grand averaged evoked potentials from Python
evokeds_exp1 <- py$evokeds_dat_exp1

# Retrive locations for the relevant electrodes from standard montage
fname_montage <- list.files(path = "renv/python", pattern = "easycap-M1.txt", recursive = TRUE, full.names = TRUE)
montage <- py$evokeds_exp1$I$Insight$ch_names %>% as_tibble() %>% rename(electrode = value) %>%
  left_join(read.delim(fname_montage) %>% as_tibble() %>% rename(electrode = Site)) %>%
  mutate(x = eegUtils:::deg2rad(Theta) * cos(eegUtils:::deg2rad(Phi)),
         y = eegUtils:::deg2rad(Theta) * sin(eegUtils:::deg2rad(Phi)))

# Define function to plot trial structure
plot_trial <- function(){
  colour_insight <- viridisLite::viridis(1, begin = 0, end = 0)
  colour_naive <- viridisLite::viridis(1, begin = 0.5, end = 0.5)
  trial_I <- ggplot() + coord_fixed() + theme_void() + xlim(c(0, 74)) + ylim(c(-5, 68)) +
    geom_text(aes(x = 37, y = 68, label = "Part I"), size = 14/.pt, family = "Helvetica", fontface = "bold") +
    geom_rect(aes(xmin = 18, xmax = 38, ymin = 13, ymax = 33), color = "black", fill = "white") +
    geom_text(aes(x = 28, y = 23, label = "+"), size = 5, family = "Helvetica", fontface = "bold") +
    geom_text(aes(x = 28, y = 10, label = "0.5 s"), color = "black", size = 10/.pt, family = "Helvetica") +
    geom_rect(aes(xmin = 36, xmax = 56, ymin = 26, ymax = 46), color = "black", fill = "white") +
    draw_image("materials/exp1/potato_masher.png", x = 36, y = 26, width = 20, height = 20) +
    geom_text(aes(x = 47, y = 23, label = "max. 3 s"), color = "black", size = 10/.pt, family = "Helvetica") +
    geom_segment(aes(x = 46, y = 13, xend = 56, yend = 20.222), arrow = arrow(length = unit(0.2, "cm")), color = "black")
  trial_II <- ggplot() + coord_fixed() + theme_void() + xlim(c(0, 74)) + ylim(c(-5, 68)) +
    geom_text(aes(x = 37, y = 68, label = "Part II"), size = 14/.pt, family = "Helvetica", fontface = "bold") +
    geom_rect(aes(xmin = 0, xmax = 20, ymin = 0, ymax = 20), color = "black", fill = "white") +
    geom_text(aes(x = 10, y = 10.5, label = "+"), size = 5, family = "Helvetica", fontface = "bold") +
    geom_text(aes(x = 10, y = -3, label = "0.5 s"), color = "black", size = 10/.pt, family = "Helvetica") +
    geom_rect(aes(xmin = 18, xmax = 38, ymin = 13, ymax = 33), color = colour_insight, fill = "white") +
    geom_text(aes(x = 28, y = 23, label = "Potato\nmashing"), color = colour_insight, size = 10/.pt, family = "Helvetica") +
    geom_text(aes(x = 28, y = 10, label = "2.5 s"), color = "black", size = 10/.pt, family = "Helvetica") +
    geom_rect(aes(xmin = 18, xmax = 38, ymin = 39, ymax = 59), color = colour_naive, fill = "white") +
    geom_text(aes(x = 28, y = 49, label = "Message\nmorsing"), color = colour_naive, size = 10/.pt, family = "Helvetica") +
    geom_text(aes(x = 28, y = 36, label = "or"), color = "black", size = 10/.pt, family = "Helvetica") +
    geom_rect(aes(xmin = 36, xmax = 56, ymin = 26, ymax = 46), color = "black", fill = "white") +
    geom_text(aes(x = 46, y = 34, label = "*"), size = 8, family = "Helvetica") +
    geom_text(aes(x = 46, y = 23, label = "0.5 s"), color = "black", size = 10/.pt, family = "Helvetica") +
    geom_rect(aes(xmin = 54, xmax = 74, ymin = 39, ymax = 59), color = "black", fill = "white") +
    draw_image("materials/exp1/potato_masher.png", x = 54, y = 39, width = 20, height = 20) +
